# R에 내장된 자료의 비구획분석 보고서 {#groupreport}

## Theoph의 보고서 {#theophgroup}

```{r code = readLines('Output-ncar/txtNCA-group-Theoph.txt'), eval = FALSE}
```

# 지원 {#support}

패키지와 관련한 모든 의문은 <shan@acp.kr> / 02-3010-4614 으로 연락 주시면 빠르게 도움 드리겠습니다.
혹은 StackOverflow^[https://stackoverflow.com]에 영어로 질문 올려주시고 링크를 보내주시면 더 좋습니다.

# sessionInfo

```{r}
devtools::session_info()
```

`r if (knitr:::is_html_output()) '# 참고문헌 {-}'`

```{r include = FALSE}
knitr::write_bib(file = "bib/packages.bib")
```

- U.S. Department of Health and Human Services Food and Drug Administration Center for Drug Evaluation and Research (CDER) , Guidance for Industry Pharmacokinetics in Pregnancy — Study Design, Data Analysis, and Impact on Dosing and Labeling, (2004)
- Malcolm Rowland and Thomas N. Tozer PharmD PhD, Clinical Pharmacokinetics and Pharmacodynamics : Concepts and Applications 4th Ed. 2010 (ISBN:9780781750097)
- Hadley Wickham, Advanced R (Chapman & Hall/CRC The R Series) 1st Ed ,25 sep 2014
- Gabrielsson J and Weiner D. Pharmacokinetic and Pharmacodynamic Data Analysis - Concepts and Applications. 5th ed. 2016. (ISBN:9198299107).
- http://www.edison.re.kr. EDISON 웹사이트. 

