# Comparison of Noncompartmental analysis software

## In this chapter

We will briefly compare NCA softwares.

## Certara Phoenix WinNonLin

<https://www.certara.com/software/pkpd-modeling-and-simulation/phoenix-winnonlin/>

### Pros

- Validated for several years
- Industry standard
- Versatile unit setting
- Easy using by GUI 
- Generating plots supported

### Cons

- Expansive (~several thousand dollars)
- Not suitable for reproducible research
- CDISC SDTM not compatible (input and output)

## R package: PKNCA

<https://github.com/billdenney/pknca>

### Pros

- Open source and free of charge
- CDISC SDTM semi compatible (output)
- Calculate partial(interval) AUC with 'linear' or 'log' interpolation method but in a cumbersome way

### Cons
 
- CDISC SDTM not compatible (input)
- More tests required
- Experience with R language required
- Generating plots not supported for now (To be supported soon)


## R package: ncar

Extension of NonCompart for CDISC input

<https://github.com/asancpt/ncar>

### Pros

- Open source and free of charge
- CDISC SDTM fully compatible (input - DM, EX, PC)
- Output of CDISC SDTM PP domain terms 
- Calculate partial(interval) AUC with 'linear' or 'log' interpolation method

### Cons

- More tests required
- Experience with R language required
- Generating plots not supported for now (To be supported soon)

## R package: ncappc

NCA Calculation and Population PK Model Diagnosis [@Acharya201683]

<https://cran.r-project.org/web/packages/ncappc/index.html>
<https://www.ncbi.nlm.nih.gov/pubmed/27000291>



```{r}
#install.packages("ncappc")
library(ncappc)
```


## Kinetica

## PK module in R

## Scientist

## PKSolver

## R package: PK

Basic Non-Compartmental Pharmacokinetics

<https://cran.r-project.org/web/packages/PK/index.html>


```{r}
#install.packages("PK")
library(PK)
```

## Summary

```{r echo = FALSE}
knitr::kable(read.csv("comparison.csv", check.names = FALSE))
```

